---
title: 'Drought prediction in continental Italy: an SpVAR approach'
output:
  html_document: default
---

We set the seed for reproducibility and import the necessary libraries and functions.

```{r package import,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
set.seed(123)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(systemfit)
  library(plm)
  library(ncdf4)
  library(abind)
  library(ggmap)
  library(Rmpfr)
  library(gridExtra)
  library(geometry)
  library(caret)
  library(cowplot)
  library(forecast)
})

setwd(
  "/Users/enrico/Library/Mobile Documents/com~apple~CloudDocs/Documents/Carriera/Scuola:Università/Bocconi Master/Thesis"
)

source("functions.r")
```

We read the SPEI data and prepare it.

```{r SPEI data read}
# Open the netCDF file with monthly SPEI data
SPEI <- nc_open("Data/spei12.nc")

# Get the data from the first variable
SPEI_data <- ncvar_get(SPEI, attributes(SPEI$var)$names[1])

# Get the longitude and latitude values
SPEI_lon <- ncvar_get(SPEI, attributes(SPEI$dim)$names[1])
SPEI_lat <- ncvar_get(SPEI, attributes(SPEI$dim)$names[2])

# Close the file
nc_close(SPEI)

# Subset the data for the relevant time steps
SPEI_last_year<-SPEI_data[,,1367:1368]
SPEI_relevant_years <- SPEI_data[, , 967:1368]
```

We read the SMoist data and prepare it.

```{r SMoist data read}
# Open the netCDF file with Soil Moisture data
SMoist <- nc_open("Data/SoilMoisture.nc")

# Get the data from the first variable
SMoist_data <- ncvar_get(SMoist, attributes(SMoist$var)$names[1])

# Order the data with the same ordering as the SPEI 
SMoist_ordered <- SMoist_data[361:720, ,]
SMoist_ordered <-
  abind(SMoist_ordered, SMoist_data[1:360, ,], along = 1)

# Get the longitude and latitude values
SMoist_lon <- ncvar_get(SMoist, attributes(SMoist$dim)$names[1])
SMoist_lat <- ncvar_get(SMoist, attributes(SMoist$dim)$names[3])

# Print a message to check if the longitudes and latitudes match with the ones from SPEI
print(paste(
  "Check on longitudes",
  toString(all(SMoist_lon - 180 == SPEI_lon)),
  "check on latitudes",
  toString(all(SMoist_lat == SPEI_lat))
))

# Close the file
nc_close(SMoist)

# Subset the data for the relevant time steps
SMoist_last_year <- SMoist_ordered[, , 659:660]
SMoist_relevant_years <- SMoist_ordered[, , 259:660]
```

We read the NDVI data and prepare it.

```{r NDVI netCDF file opening}
# Open the netCDF file with NDVI data
NDVI_read <- nc_open("Data/NDVI.nc")

# Get the data from the first variable
NDVI <- ncvar_get(NDVI_read, attributes(NDVI_read$var)$names[1])

# Get the longitude and latitude values
NDVI_lon <- ncvar_get(NDVI_read, attributes(NDVI_read$dim)$names[1])
NDVI_lat <- ncvar_get(NDVI_read, attributes(NDVI_read$dim)$names[2])

# Print a message to check if the longitudes and latitudes match with the ones from SPEI
print(paste(
  "Check on longitudes",
  all(NDVI_lon == SPEI_lon),
  "check on latitudes",
  all(NDVI_lat == SPEI_lat)
))

# Close the file
nc_close(NDVI_read)

# Subset the data for the relevant time steps
NDVI_last_year <- NDVI[, , 401:402]
```

We read the PET data and prepare it.

```{r PET}
# Open the first netCDF file with monthly PET data and store it
PET <- nc_open("Data/cru_ts4.03.1981.1990.pet.dat.nc")
PET_data <- ncvar_get(PET, attributes(PET$var)$names[1])

# Subset the data for the relevant time steps
PET_data <- PET_data[, , 7:120]

# Close the file
nc_close(PET)

# Open the second netCDF file with monthly PET data and store it
PET <- nc_open("Data/cru_ts4.03.1991.2000.pet.dat.nc")
PET_data_2 <- ncvar_get(PET, attributes(PET$var)$names[1])

# Close the file
nc_close(PET)

# Open the third netCDF file with monthly PET data and store it
PET <- nc_open("Data/cru_ts4.03.2001.2010.pet.dat.nc")
PET_data_3 <- ncvar_get(PET, attributes(PET$var)$names[1])

# Close the file
nc_close(PET)

# Open the fourth netCDF file with monthly PET data and store it
PET <- nc_open("Data/cru_ts4.03.2011.2018.pet.dat.nc")
PET_data_4 <- ncvar_get(PET, attributes(PET$var)$names[1])

# Subset the data for the relevant time steps
PET_data_4 <- PET_data_4[, , 1:48]

# Get the longitude and latitude values
PET_lon <- ncvar_get(PET, attributes(PET$dim)$names[1])
PET_lat <- ncvar_get(PET, attributes(PET$dim)$names[2])

# Print a message to check if the longitudes and latitudes match with the ones from SPEI
print(paste(
  "Check on longitudes",
  all(PET_lon == SPEI_lon),
  "check on latitudes",
  all(PET_lat == SPEI_lat)
))

# Close the file
nc_close(PET)

# Merge the 4 PET datasets together
PET <- array(0, dim = c(720, 360, 402))
PET[, , ] <- c(PET_data, PET_data_2, PET_data_3, PET_data_4)
```

We read the PRE (Monthly Precipitation) data and prepare it.

```{r PRE}
# Open the first netCDF file with monthly Precipitation dataand store it
PRE <- nc_open("Data/cru_ts4.03.1981.1990.pre.dat.nc")
PRE_data <- ncvar_get(PRE, attributes(PRE$var)$names[1])

# Subset the data for the relevant time steps 
PRE_data<-PRE_data[,,7:120]

# Close the file
nc_close(PRE)

# Open the second netCDF file with monthly Precipitation data and store it
PRE<-nc_open("Data/cru_ts4.03.1991.2000.pre.dat.nc")
PRE_data_2 <- ncvar_get(PRE, attributes(PRE$var)$names[1])

# Close the file
nc_close(PRE)

# Open the third netCDF file with monthly Precipitation data and store it
PRE<-nc_open("Data/cru_ts4.03.2001.2010.pre.dat.nc")
PRE_data_3 <- ncvar_get(PRE, attributes(PRE$var)$names[1])

# Close the file
nc_close(PRE)

# Open the fourth netCDF file with monthly Precipitation data and store it
PRE<-nc_open("Data/cru_ts4.03.2011.2018.pre.dat.nc")
PRE_data_4 <- ncvar_get(PRE, attributes(PRE$var)$names[1])

# Subset the data for the relevant time steps
PRE_data_4<-PRE_data_4[,,1:48]

# Get the longitude and latitude values
PRE_lon <- ncvar_get(PRE, attributes(PRE$dim)$names[1])
PRE_lat <- ncvar_get(PRE, attributes(PRE$dim)$names[2])

# Print a message to check if the longitudes and latitudes match with the ones from SPEI
print(paste(
  "Check on longitudes",
  all(PRE_lon == SPEI_lon),
  "check on latitudes",
  all(PRE_lat == SPEI_lat)
))

# Close the file
nc_close(PRE)

# Merge the 4 Precipitation datasets together
PRE<-array(0, dim=c(720,360,402))
PRE[,,]<-c(PRE_data, PRE_data_2,PRE_data_3,PRE_data_4)

```

We clean-up the workspace by removing unnecessary objects.

```{r clean-up}
# Remove the variables SPEI and SMoist from the workspace
rm(SPEI, SMoist)

# Rename the variables SPEI and Smoist variables
SPEI <- SPEI_relevant_years
SMoist <- SMoist_relevant_years

# Assign the variables lats and lons to the latitude and longitude of SPEI
lats <- SPEI_lat
lons <- SPEI_lon

# Remove variables that won't be used again from the workspace
rm(
  SPEI_relevant_years,
  SMoist_relevant_years,
  SPEI_lat,
  SMoist_lat,
  SPEI_lon,
  SMoist_lon,
  SMoist_ordered,
  NDVI_lon,
  NDVI_lat,
  SPEI_data,
  SMoist_data,
  NDVI_read,
  PET_lat,
  PET_lon,
  PRE_lat,
  PRE_lon,
  PET_data,
  PET_data_2,
  PET_data_3,
  PET_data_4,
  PRE_data,
  PRE_data_2,
  PRE_data_3,
  PRE_data_4
)
```

We create arrays containing the latitudes and longitudes of the region (Continental Italy) we'll analyze.

```{r Latitudes long}
reg_latitudes <-
  c(
    46.75,
    46.25,
    46.25,
    45.75,
    45.25,
    44.75,
    44.25,
    43.75,
    43.75,
    43.25,
    42.75,
    42.25,
    41.75,
    41.25,
    40.75,
    40.25,
    39.75,
    39.25,
    38.75,
    38.25
  )
reg_long_min <-
  c(
    10.25,
    8.25,
    9.25,
    6.75,
    6.75,
    6.75,
    6.75,
    7.25,
    10.25,
    10.25,
    10.75,
    11.75,
    12.25,
    13.25,
    14.25,
    15.25,
    15.75,
    16.25,
    16.25,
    15.75
  )

reg_long_max <-
  c(
    12.75,
    8.25,
    13.75,
    13.75,
    12.25,
    12.25,
    12.25,
    7.75,
    13.75,
    13.75,
    14.25,
    14.25,
    15.75,
    16.75,
    17.75,
    18.25,
    16.25,
    16.75,
    16.75,
    16.25
  )
```

We now scale the data and then format it as needed.

```{r data preparation}
test_df <- create_test_df(reg_latitudes,
                          reg_long_min,
                          reg_long_max,
                          lats,
                          lons,
                          SPEI,
                          SMoist,
                          NDVI)

#We store the scaling factors for future re-use
scale_SMoist <- sd(test_df$SMoist)
scale_NDVI <- sd(test_df$NDVI, na.rm = TRUE)
#Since we don't scale the SPEI value we set the scale factor to 1
scale_SPEI <- 1

#We scale the data
test_df$SMoist <- (test_df$SMoist) / (scale_SMoist)
test_df$NDVI <- (test_df$NDVI) / (scale_NDVI)

#We create a dataframe for each of the variables
test_df_spei <-
  data.frame(split(test_df$SPEI, as.factor(test_df$index)))
test_df_ndvi <-
  data.frame(split(test_df$NDVI, as.factor(test_df$index)))
test_df_smoist <-
  data.frame(split(test_df$SMoist, as.factor(test_df$index)))
```

We create two other objects needed in the future analysis (the regions indexes and the weights matrix).

```{r other objects creation}
#index_pairs contains the Longitudes and Latitudes indexes of all the regions under consideration
index_pairs <- unique(test_df$index)

#weights is the connectivity matrix
weights <- prepare_weights(index_pairs, lats, lons)
```

# Unit Root testing

We now perform unit root testing as described in Section 3.3. We first look at the estimated $\lambda$ for the SPEI variable on our data.

```{r lambda spei}
compute_lambda(test_df_spei, weights)
```

We then look at the estimated $\lambda$ for the SMoist variable.

```{r lambda smoist}
compute_lambda(test_df_smoist, weights)
```

We also look at the estimated $\lambda$ for the NDVI variable.

```{r lambda ndvi}
compute_lambda(test_df_ndvi, weights)
```

Finally, we perform a Monte Carlo simulation to obtain the critical values of our test.

```{r lambda mc, cache=TRUE}
lambda_mc <-
  mc_critical_values(10000, 139, 500, weights, 3061792, FALSE)
```

# Model fitting

We fit our model using SUR (implemented in the systemfit package) after having prepared the data for it.

```{r SpVAR fit,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
spvar_df <-
  prepare_spvar_df(test_df, test_df_spei, test_df_smoist, test_df_ndvi, lags =
                     1)


spvar_fit <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist
    ),
    data = spvar_df,
    method = "SUR"
  )
```

We also fit a model with two temporal lags to compare them.

```{r 2 lagged SpVAR fit,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
spvar_df_2 <-
  prepare_spvar_df(test_df, test_df_spei, test_df_smoist, test_df_ndvi, lags =
                     2)

spvar_fit_2 <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist
    ),
    data = spvar_df_2,
    method = "SUR"
  )
```

To make this comparison and choose which one to use we use BIC.

```{r Compare SpVAR fit}
print(paste( "The BIC of the first model is", BIC(spvar_fit)))
print(paste( "The BIC of the second model is", BIC(spvar_fit_2)))
```

Since the BIC of the first model is substantially lower we choose it. We now look at the coefficients of this chosen model.

```{r SpVAR coefficients, warning=FALSE}
summary(spvar_fit)
```

## Predictive performance

To evaluate the predictive capabilities of our model, we fit it on the first 300 months of data. We also fit 2 simpler models (VAR and SAR) with which we will compare ours. More details on this are in Section 4.2.

```{r short term models fitting,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
# Prepare the dataset containing only the first 300 months
df_short <-
  prepare_spvar_df(test_df, test_df_spei, test_df_smoist, test_df_ndvi, T =
                     301)

spvar_fit_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist
    ),
    data = df_short,
    method = "SUR"
  )

var_fit_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_smoist -sp_ndvi,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_smoist -sp_ndvi,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_smoist -sp_ndvi
    ),
    data = df_short,
    method = "SUR"
  )

sar_fit_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist -sp_smoist -sp_ndvi -ndvi_lag -smoist_lag,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist -sp_smoist -sp_spei -spei_lag -smoist_lag,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_ndvi -ndvi_lag -spei_lag
    ),
    data = df_short,
    method = "SUR"
  )
```

We now look at the adjusted R\^2 of the three models for the SPEI equation.

```{r adj r squared}
print(paste("The adjusted R Squared of the SpVAR is:",adj_r2(spvar_fit_short$eq[[1]])))
print(paste("The adjusted R Squared of the VAR is:",adj_r2(var_fit_short$eq[[1]])))
print(paste("The adjusted R Squared of the SAR is:",adj_r2(sar_fit_short$eq[[1]])))
```

We subset the data as needed for making predictions and set all the missing values to 0 so that they don't impact our predictions.

```{r missing to 0}
data_pred_short <- spvar_df[((299 * 139) + 1):dim(spvar_df)[1],]

data_pred_short[is.na(data_pred_short)] = 0
```

We now look at predictive performances of the three models using the $SMAPE$ and the $RMSE$.
We start by comparing the error rates using the three models for predictions up to a year in advance

```{r up to 12 months prediction comparison, cache=TRUE}
for (i in c(1, 3, 6, 12)) {
  print(paste(i, "steps ahead:"))
  spvar_err <-
    compute_pred_steps(102 - i, i, spvar_fit_short, data_pred_short)
  print(paste("The SMAPE and the RMSE are as follows for the SpVAR:", spvar_err[[2]],spvar_err[[1]]))
  var_err <-
    compute_pred_steps(102 - i, i, var_fit_short, data_pred_short, i+1)
  print(paste("The SMAPE and the RMSE are as follows for the VAR:", var_err[[2]],var_err[[1]] ))
  sar_err <-
    compute_pred_steps(102 - i, i, sar_fit_short, data_pred_short, i)
  print(paste("The SMAPE and the RMSE are as follows for the SAR:", sar_err[[2]], sar_err[[1]]))
}
```

We check if the differences are significant using the Diebold-Mariano test.

```{r dm test, cache=TRUE}
for (i in c(1, 3, 6, 12)) {
  spvar_res <-
    compute_pred_steps(102 - i,
                       i,
                       spvar_fit_short,
                       data_pred_short,
                       value = "res")
  sar_res <-
    compute_pred_steps(102 - i,
                       i,
                       sar_fit_short,
                       data_pred_short,
                       i,
                       value = "res")
  var_res <-
    compute_pred_steps(102 - i,
                       i,
                       var_fit_short,
                       data_pred_short,
                       i + 1,
                       value = "res")
  spvar_res <- spvar_res[2:((102 - i) * 44)]
  sar_res <- sar_res[2:((102 - i) * 44)]
  var_res <- var_res[2:((102 - i) * 44)]
  print(dm.test(spvar_res, sar_res, h = i))
  print(dm.test(spvar_res, var_res, h = i))
}
```

We now fit the models on a shorter data-set (containing only the first 150 months) to see the performance on longer term predictions (60 and 120-months ahead).

```{r fit shorter models,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
# Prepare the dataset containing only the first 150 months
df_super_short <-
  prepare_spvar_df(test_df, test_df_spei, test_df_smoist, test_df_ndvi, T =
                     151)


spvar_fit_super_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist
      ),
    data = df_super_short,
    method = "SUR"
  )


var_fit_super_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_ndvi -sp_smoist,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_ndvi -sp_smoist,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_ndvi -sp_smoist
    ),
    data = df_super_short,
    method = "SUR"
  )

sar_fit_super_short <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . -spei -ndvi -smoist -sp_smoist -sp_ndvi -ndvi_lag -smoist_lag,
      eqndvi = ndvi ~ 0 + . -spei -ndvi -smoist -sp_smoist -sp_spei -spei_lag -smoist_lag,
      eqsmoist = smoist ~ 0 + . -spei -ndvi -smoist -sp_spei -sp_ndvi -ndvi_lag -spei_lag
    ),
    data = df_super_short,
    method = "SUR"
  )
```

We again subset the data as needed and set all the missing values to 0 so that they don’t impact our predictions.

```{r missing to 0 short}
data_pred_super_short <- spvar_df[((149 * 139) + 1):dim(spvar_df)[1],]

data_pred_super_short[is.na(data_pred_super_short)] = 0
```

Finally, we check the performances on these longer-term predictions and check if the differences between models are significant with the Diebold-Mariano test.

```{r long-term predictions, cache=TRUE}
for (i in c(60, 120)) {
  print(paste(i, "steps ahead:"))
  spvar_err <-
    compute_pred_steps(252 - i, i, spvar_fit_super_short, data_pred_super_short)
  print(paste(
    "The SMAPE and RMSE are as follows for the SpVar",
    spvar_err[[2]],
    spvar_err[[1]]
  ))
  var_err <-
    compute_pred_steps(252 - i, i, var_fit_super_short, data_pred_super_short, i+1)
  print(paste(
    "The SMAPE and RMSE are as follows for the VAR",
    var_err[[2]],
    var_err[[1]]
  ))
  sar_err <-
    compute_pred_steps(252 - i, i, sar_fit_super_short, data_pred_super_short, i)
  print(paste(
    "The SMAPE and RMSE are as follows for the SAR",
    sar_err[[2]],
    sar_err[[1]]
  ))
  spvar_res <-
    compute_pred_steps(252 - i,
                       i,
                       spvar_fit_super_short,
                       data_pred_super_short,
                       value = "res")
  sar_res <-
    compute_pred_steps(252 - i,
                       i,
                       sar_fit_super_short,
                       data_pred_super_short,
                       i,
                       value = "res")
  var_res <-
    compute_pred_steps(252 - i,
                       i,
                       var_fit_super_short,
                       data_pred_super_short,
                       i + 1,
                       value = "res")
  spvar_res <- spvar_res[2:((252 - i) * 139)]
  sar_res <- sar_res[2:((252 - i) * 139)]
  var_res <- var_res[2:((252 - i) * 139)]
  print(dm.test(spvar_res, sar_res), h = i)
  print(dm.test(spvar_res, var_res), h = i)
}

```

# Impulse Response Functions

We now look at impulse response functions starting from same region IRFs.

```{r northeast, fig.height =6, fig.width =8, cache=TRUE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache=TRUE}
reg_2 <- 106
t_irf <- 100

quantiles <- bootstrap(t_irf, reg_2, spvar_fit, spvar_df, 1000)
multiplot_irf_sameregion(spvar_fit, spvar_df, reg_2, 100, quantiles)
```

We now look at the IRF in neighbouring regions when shocking a region having transitional climate.

```{r north east map, fig.height =10, fig.width =9, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
plot <- multiplot_irf_all(spvar_fit, 100, reg_2, plot_all = FALSE)

map <- multiplot_irf_map(spvar_fit, index_pairs, test_df, 50, reg_2, 20)
map[[1]] <-
  map[[1]] + theme(plot.margin = unit(c(
    b = 0.3,
    l = 0,
    t = 0.3,
    r = 0
  ), "cm"))
map[[2]] <-
  map[[2]] + theme(plot.margin = unit(c(
    b = 0.3,
    l = 0,
    t = 0.3,
    r = 0
  ), "cm"))
map[[3]] <-
  map[[3]] + theme(plot.margin = unit(c(
    b = 0.3,
    l = 0,
    t = 0.3,
    r = 0
  ), "cm"))


grid.arrange(
  plot[[1]],
  map[[1]],
  plot[[2]],
  map[[2]],
  plot[[3]],
  map[[3]],
  nrow = 3,
  ncol = 2,
  widths = c(1, 1.1)
)
```

Finally, we plot the IRFs for neighbouring regions when shocking regions with alpine and Mediterranean climate. 

```{r south, fig.height =8, fig.width =8}
reg_1 <- 19
reg_3 <- 137

plot_1 <- multiplot_irf_all(spvar_fit, 75, reg_1, plot_all = FALSE)
plot_2 <- multiplot_irf_all(spvar_fit, 75, reg_3, plot_all = FALSE)

grid.arrange(plot_1[[1]], plot_2[[1]],   plot_1[[2]], plot_2[[2]],
             plot_1[[3]], plot_2[[3]],
             ncol = 2)
```


# Other plots

The following plots the mean monthly precipitations and the average PET for each region.

```{r, fig.width=8, fig.height=3, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache = TRUE}
# Prepare a dataframe for plotting
plot_df <-
  create_plot_df(1,
                 reg_latitudes ,
                 reg_long_min,
                 reg_long_max,
                 lats,
                 lons,
                 SPEI,
                 NDVI,
                 SMoist)

# Compute the mean precipitation and PET and store them in the dataframe
for (i in 1:139) {
  lon_ind <- as.numeric(strsplit(index_pairs[i], ",")[[1]][2])
  lat_ind <- as.numeric(strsplit(index_pairs[i], ",")[[1]][1])
  plot_df[plot_df$lat_ind == lat_ind &
            plot_df$lon_ind == lon_ind, 5] <-
    mean(PET[lon_ind, lat_ind, 6:402])
  plot_df[plot_df$lat_ind == lat_ind &
            plot_df$lon_ind == lon_ind, 6] <-
    mean(PRE[lon_ind, lat_ind, 6:402])
}

#Duplicate the dataframe
plot_df_2 <- plot_df

#Rename a column for plotting
colnames(plot_df_2)[5] <- "weights"
colnames(plot_df)[6] <- "weights"

plot_1 <-
  create_plot(plot_df_2,
              var = "Mean PET", scale_fixed = TRUE)

plot_2 <-
  create_plot(plot_df,
              var = "Mean PRE", scale_fixed = TRUE)


grid.arrange(plot_1, plot_2, ncol = 2)
```

The following plots the evolution of PET and PRE (monthly precipitations) in certain regions (one for each climatic area) to highlight the differences between them.

```{r, fig.width=8, fig.height=4, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache = TRUE}
# Prepare a dataframe with the data
test_df_pet <- create_test_df(reg_latitudes,
                              reg_long_min,
                              reg_long_max,
                              lats,
                              lons,
                              PET,
                              PRE,
                              SPEI)

# Rename relevant columns
names(test_df_pet)[6] <- "PET"
names(test_df_pet)[7] <- "PRE"

plot_1_pre <-
  ggplot(data = test_df_pet[test_df_pet$index == index_pairs[reg_1] |
                              test_df_pet$index == index_pairs[reg_2] |
                              test_df_pet$index == index_pairs[reg_3], ]) +
  geom_smooth(aes(x = month, y = PET, color = index)) +
  guides(colour = "none") +
  scale_color_manual(
    name = "Climate",
    labels = c("Mediterranean", "Transitional", "Alpine"),
    values = c("274,384",  "271,384", "262,392")
  ) + labs(x = "Months")


plot_2_pre <-
  ggplot(data = test_df_pet[test_df_pet$index == index_pairs[reg_1] |
                              test_df_pet$index == index_pairs[reg_2] |
                              test_df_pet$index == index_pairs[reg_3],]) +
  geom_smooth(aes(x = month, y = PRE, color = index)) + scale_color_manual(
    name = "Climate",
    labels = c("Mediterranean", "Transitional", "Alpine"),
    values = c("274,384", "271,384", "262,392")
  ) + labs(x = "Months")

grid.arrange(plot_1_pre, plot_2_pre, ncol = 2, widths = c(0.9, 1.3))
```

The following plots the region under consideration (Continental Italy) and how we divide it to exemplify some concepts of spatial analysis together with the weighting scheme we use.

```{r Regions plot, fig.height =4, fig.width =8, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache=TRUE}
# Prepare a plot showcasing the region under consideration
plot_df<-create_plot_df(1,reg_latitudes, reg_long_min, reg_long_max, lats, lons, SPEI, SMoist, NDVI)
reg_plot <- qmplot(
  y = lat_plot,
  x = lon_plot,
  data = plot_df,
  geom = "blank",
  maptype = "toner-background",
  zoom = 7
) +
  geom_rect(
    aes(
      ymin = lat_min,
      ymax = lat_max,
      xmin = lon_min,
      xmax = lon_max,
      alpha = 0.01,
      color = "white",
      fill = "grey"
    )
  ) + scale_alpha_continuous(guide = "none") +
  scale_color_discrete(guide = "none", type = "white") +
  scale_fill_discrete(guide = "none", type = "grey") +
  annotate(
    "text",
    x = 12.75,
    y = 46.75,
    label = "B",
    size = 4,
    color = "black"
  ) +
  annotate(
    "text",
    x = 11.75,
    y = 45.25,
    label = "A",
    size = 4,
    color = "black"
  ) +
  annotate(
    "text",
    x = 15.75,
    y = 40.75,
    label = "C",
    size = 4,
    color = "black"
  )

# Prepare a plot showcasing the weighting scheme
plot_weights <- create_plot_df_weights(weights[reg_2,])
weight_plot <- create_plot(plot_weights, var = "Weights")
reg_plot <-
  reg_plot + theme(plot.margin = unit(c(0, 0.75, 0, 0), "cm"))
grid.arrange(reg_plot,
             weight_plot,
             ncol = 2,
             widths = c(0.87, 1))
```

The following plots the distribution of $\lambda$ under the first and tenth unit root test

```{r lambda mc plot, fig.height =4, fig.width =8, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache = TRUE}
lambda_first_test <- tibble(x = lambda_mc[1:10000, ])
lambda_tenth_test <- tibble(x = lambda_mc[90001:100000, ])

hist1 <-
  ggplot(data = lambda_first_test) + geom_histogram(
    aes(x = x),
    binwidth = 0.0005,
    color = "black",
    fill = "black"
  ) + labs(x = "Lambda in test 1", y = "Count") +
  geom_vline(aes(xintercept = quantile(x, probs = 0.05)), color = "red") +
  scale_y_continuous(limits = c(0, 570))

hist2 <-
  ggplot(data = lambda_tenth_test) + geom_histogram(
    aes(x = x),
    binwidth = 0.0005,
    color = "black",
    fill = "black"
  ) + labs(x = "Lambda in test 10", y = NULL) +
  scale_y_continuous(limits = c(0, 570), guide = "none") +
  geom_vline(aes(xintercept = quantile(x, probs = 0.05)), color = "red")


grid.arrange(hist1, hist2, ncol = 2, widths = c(1.2, 1))
```

The following plots the number of droughts of each type in the 402 months under consideration.

```{r, fig.width=8, fig.height=3, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache=TRUE}
plots_droughts <-
  long_term_prediction_plot(
    test_df_spei[1:402,], plot="initial"
  )
```

The following plots the predicted number of droughts in the second 201 periods of the period under consideration (below) against the actual number of them in the data (above). Note that we use a model fit only on the first 201 months.

```{r, fig.width=8, fig.height=6, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache=TRUE}
# Prepare the data
test_df_201<- test_df[((139 * 199) + 1):(139 * 201), ]
test_df_201$month <- c(rep(1, times = 139), rep(2, times = 139))
test_df_201_spei <- test_df_spei[200:201, ]
test_df_201_ndvi <- test_df_ndvi[200:201, ]
test_df_201_smoist <- test_df_smoist[200:201, ]

# Fit a model on the first 201 periods
spvar_df_201 <- spvar_df[1:(201 * 139), ]
spvar_fit_201 <-
  systemfit(
    c(
      eqspei = spei ~ 0 + . - spei - ndvi - smoist,
      eqndvi = ndvi ~ 0 + . - spei - ndvi - smoist,
      eqsmoist = smoist ~ 0 + . - spei - ndvi - smoist
    ),
    data = spvar_df_201,
    method = "SUR"
  )

# Create the plot
plots_droughts_201 <-
  long_term_prediction_plot(
    test_df_spei[201:402, ],
    test_df_201,
    test_df_201_spei,
    test_df_201_smoist,
    test_df_201_ndvi,
    t_pred = 201,
    spvar_fit = spvar_fit_201
  )
```

The following plots the predicted number of droughts in the next 402 periods (the 402 months after those covered by the dataset, i.e. starting from January 2014).

```{r, fig.width=8, fig.height=3, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all', cache=TRUE}
# Create a table with the data of the last 2 years
test_df_last_year <- create_test_df(
  reg_latitudes,
  reg_long_min,
  reg_long_max,
  lats,
  lons,
  SPEI_last_year,
  SMoist_last_year,
  NDVI_last_year
)

# Subset the data
test_df_last_year <- test_df_last_year[1:278, ]

#Create a separate df for each variable
test_df_last_year_spei <-
  data.frame(split(test_df_last_year$SPEI, as.factor(test_df_last_year$index)))
test_df_last_year_ndvi <-
  data.frame(split(test_df_last_year$NDVI, as.factor(test_df_last_year$index)))
test_df_last_year_smoist <-
  data.frame(split(
    test_df_last_year$SMoist,
    as.factor(test_df_last_year$index)
  ))


plots_droughts_future <-
  long_term_prediction_plot(
    test_df_spei[201:402, ],
    test_df_last_year,
    test_df_last_year_spei,
    test_df_last_year_smoist,
    test_df_last_year_ndvi,
    t_pred = 402,
    spvar_fit = spvar_fit,
    plot = "future"
  )
```



